---
title: "Germline-somatic interaction - Figure 2"
author:
  - name: Anton Safonov, Luca Boscolo Bielo, David N Brown
output:
  html_document:
    theme: cerulean
date: "`r format(Sys.time(), '%d %B, %Y')`"
---

```{r, eval = TRUE, echo = FALSE, results = 'hide', include = FALSE}
library('BSgenome.Hsapiens.UCSC.hg19')
library('ComplexHeatmap')
library('copynumber')
library('data.table')
library('doMC')
library('dplyr')
library('exact2x2')
library('foreach')
library('fuzzyjoin')
library('gdata')
library('GenomicRanges')
library('ggforce')
library('ggnomics')
library('ggplot2')
library('gplots')
library('ggrepel')
library('ggsignif')
library('gridExtra')
library('Gviz')
library('htmltools')
library('jsonlite')
library('knitr')
library('logistf')
library('lubridate')
library('maftools')
library('magrittr')
library('mime')
library('pander')
library('parallel')
library('RColorBrewer')
library('readr')
library('reshape2')
library('superheat')
library('survival')
library('survminer')
library('tidyr')
library('tidyverse')
library('viridis')

registerDoMC(8)

'%notin%' <- function(x,y) {
	!(x %in% y)
}

'colMax' <- function(data) {
	sapply(data, max, na.rm = TRUE)
}

'numextract' <- function(string) {
	str_extract(string, "\\-*\\d+\\.*\\d*")
}

'tabf' <- function(x, y, i) {
	z = c(colnames(x)[i+4], 
	      exp(y$coefficients[[2]]),
	      exp(y$ci.lower[[2]]),
	      exp(y$ci.upper[[2]]),
	      y$prob[[2]],
	      sum(x[, i+4])/dim(x)[1]*100,
	      sum(x[x$GL == 1,i+4])/dim(x[x$GL == 1, ])[1]*100,
	      sum(x[x$GL == 1,i+4]),
	      exp(logistf(GL ~ x[,i+4], data=x)$coefficients[[2]]),
	      logistf(GL ~ x[,i+4], data=x)$prob[[2]])
	return(invisible(z))
}
```
<br/>

```{r, eval = TRUE, echo = FALSE, message = FALSE, include = TRUE}
cdk_b2_all = suppressWarnings(readr::read_csv(file = "../data/Supp_Fig2ab_CDK46_PFS_ANON.csv.gz", col_names = TRUE, col_types = cols(.default = col_character()))) %>%
	     readr::type_convert() %>%
	     dplyr::select(-X1) %>%
	     dplyr::distinct() %>%
	     dplyr::rename_all( function(x) { c("Patient.ID", "Tx.Line", "time.months", "event", "ET_partner", "B2", "cohort", "RB1_LOH") }) %>%
	     dplyr::mutate(Tx.Line = as.numeric(Tx.Line))

cdk_b2_1 = cdk_b2_all %>%
	   dplyr::filter(Tx.Line == 1)

cdk_b2_all = cdk_b2_all %>%
	     dplyr::mutate(Tx.Line.c = case_when(
		     	Tx.Line == 1 ~ "1",
		       (Tx.Line > 1 & Tx.Line < 4) ~ "2to3",
		       (Tx.Line > 3 & Tx.Line < 6) ~ "4to5",
		       Tx.Line > 5 ~ "6+"
	     )) %>%
	     dplyr::mutate(Tx.Line.c = factor(Tx.Line.c, levels = c("1", "2to3", "4to5", "6+")))

cdk_b2.f1.plot = survfit(Surv(time.months, event) ~ B2, data = cdk_b2_1, conf.type = "log-log") %>%
		 ggsurvplot(palette = c("#00B26E", "#F54782"),
			    risk.table = TRUE,
			    tables.height = 0.20,
			    tables.theme = theme_cleantable(),
			    censor = TRUE,
			    break.time.by = 6,
			    fontsize = 3,
			    axes.offset = FALSE,
			    xlab = "Follow up Time (mo)",
			    xlim = c(0,78))

cdk_b2.fall.plot = survfit(Surv(time.months, event) ~ B2, data = cdk_b2_all, conf.type = "log-log") %>%
		   ggsurvplot(palette = c("#00B26E", "#F54782"),
			      risk.table = TRUE,
			      tables.height = 0.20,
			      risk.table.font.size = 10,
			      tables.theme = theme_cleantable(),
			      censor = TRUE,
			      break.time.by = 6,
			      fontsize = 3,
			      axes.offset = FALSE,
			      xlab = "Follow up Time (mo)",
			      xlim = c(0,78))

```{r figure_2a, eval = TRUE, echo = FALSE, message = FALSE, include = TRUE, fig.align = 'left', fig.height = 5, fig.width = 5.5, fig.cap = '**Figure 2a:**'}
print(cdk_b2.f1.plot)
```
<br/>

```{r figure_2b, eval = TRUE, echo = FALSE, message = FALSE, include = TRUE, fig.align = 'left', fig.height = 5, fig.width = 5.5, fig.cap = '**Figure 2b:**'}
print(cdk_b2.fall.plot)
```

<br/>

```{r, eval = TRUE, echo = FALSE, message = FALSE, include = TRUE}
summary(coxph(Surv(time.months, event) ~ B2 + ET_partner, data = cdk_b2_1))
summary(coxph(Surv(time.months, event) ~ B2 + ET_partner + Tx.Line.c, data = cdk_b2_all))
```

```{r figure_2d, eval = TRUE, echo = FALSE, message = FALSE, include = TRUE, fig.align = 'left', fig.height = 7, fig.width = 5.5, fig.cap = '**Figure 2d:**'}
all.lines = readr::read_csv(file = "../data/Supp_Fig2d_swimmer_ANON.csv.gz", col_names = TRUE, col_types = cols(.default = col_character())) %>%
	    readr::type_convert() %>%
	    dplyr::rename(Reason.Stop = Reason_Stop,
			  Best.Response = Best_Resposne,
			  Start.Date.norm = Deidentified_Start_Date,
			  End.Date.norm = Deidentified_End_Date)

plot_ = all.lines %>%
	ggplot(mapping = aes(y = index, fill = Tx.Type)) +
	scale_fill_manual(values = c("cdk" = "#4271D6",
				     "parp" = "#00B26E",
				     "bet" = "#F4F4F4")
	) +
	xlab("Time (months)") +
	ylab("") + 
	geom_rect(
		aes(xmin = Start.Date.norm,
		    xmax = End.Date.norm,
		    ymin = index - 0.3,
		    ymax = index + 0.3,
		    fill = Tx.Type),
		size = 0.2, color = "black"
	) +
	geom_point(
		data = subset(all.lines, Reason.Stop == "Disease Progression" & Tx.Type %in% c("cdk", "parp")),
		aes(x = End.Date.norm,
		    y = index),
		color = "black",
		shape = 18,
		size = 3.0
	) +
	geom_point(
		data = subset(all.lines, Reason.Stop %in% c("MD Preference", "Toxicity") & Tx.Type %in% c("cdk", "parp")),
		aes(x = End.Date.norm,
		    y = index),
		color = "#BBBBBC",
		shape = 18,
		size = 3.0
	) +
	geom_segment(
		data = subset(all.lines, Reason.Stop == "Ongoing" & Tx.Type %in% c("cdk", "parp")),
		aes(x = End.Date.norm,
		    y = index,
		    xend = End.Date.norm + 2,
		    yend = index),
		arrow = arrow(length = unit(0.2, "cm"))
	) +
	geom_point(
		data = subset(all.lines, Rapid.Prog == 0 & Best.Response == "CR"),
		aes(x = Scan.Date.Norm,
		    y = index),
		color = "#FFFDEC",
		shape = 19,
		size = 3
	) +
	geom_point(
		data = subset(all.lines, Rapid.Prog == 0 & Best.Response == "PR"),
		aes(x = Scan.Date.Norm,
		    y = index),
		color = "#9cc0cd",
		shape = 19,
		size = 3
	) +
	geom_point(
		data = subset(all.lines, Rapid.Prog == 0 & Best.Response == "SD"),
		aes(x = Scan.Date.Norm,
		    y = index),
		color = "#6b8590",
		shape = 19,
		size = 3
	) +
	# Tx Line CDK
	geom_point(
		data = subset(all.lines, Tx.Type == "cdk" & Tx.Line == 1),
		aes(x = -2,
		    y = index),
		fill = "#E2EEF5",
		color = "black",
		shape = 22,
		size = 4.5
	) +
	geom_point(
		data = subset(all.lines, Tx.Type == "cdk" & Tx.Line == 2),
		aes(x = -2,
		    y = index),
		fill = "#8FB9D1",
		color = "black",
		shape = 22,
		size = 4.5
	) +
	geom_point(
		data = subset(all.lines, Tx.Type == "cdk" & Tx.Line > 2),
		aes(x = -2,
		    y = index),
		fill = "#2389AF",
		color = "black",
		shape = 22,
		size = 4.5
	) +
	geom_point(
		data = subset(all.lines, Tx.Type == "cdk" & LOH_status == "AI_LOH_ALT"),
		aes(x = -4,
		    y = index),
		fill = "#DF3E9C",
		color = "black",
		shape = 22,
		size = 4.5
	) +
	geom_point(
		data = subset(all.lines, Tx.Type == "cdk" & LOH_status == "No_LOH"),
		aes(x = -4,
		    y = index),
		fill = "#FAC1E3",
		color = "black",
		shape = 22,
		size = 4.5
	) +
	geom_point(
		data = subset(all.lines, Tx.Type == "cdk" & LOH_status == "Non_evaluable"),
		aes(x = -4,
		    y = index),
		fill = "#FFFFFF",
		color = "black",
		shape = 22,
		size = 4.5
	) +
	geom_point(
		data = subset(all.lines, Tx.Type == "cdk" & CDK == "Palbociclib"),
		aes(x = -6,
		    y = index),
		fill = "#EEDCD6",
		color = "black",
		shape = 22,
		size = 4.5
	) +
	geom_point(
		data = subset(all.lines, Tx.Type == "cdk" & CDK == "Abemaciclib"),
		aes(x = -6,
		    y = index),
		fill = "#D38D7B",
		color = "black",
		shape = 22,
		size = 4.5
	) +
	geom_point(
		data = subset(all.lines, Tx.Type == "cdk" & CDK == "Ribociclib"),
		aes(x = -6,
		    y = index),
		fill = "#955442",
		color = "black",
		shape = 22,
		size = 4.5
	) +
  	geom_point(
		data = subset(all.lines, Tx.Type == "cdk" & AI.SERD == "AI"),
		aes(x = -8,
		    y = index),
		fill = "#E9CBF2",
		color = "black",
		shape = 22,
		size = 4.5
	) +
	geom_point(
		data = subset(all.lines, Tx.Type == "cdk" & AI.SERD == "SERD"),
		aes(x = -8,
		    y = index),
		fill = "#BC59D3",
		color = "black",
		shape = 22,
		size = 4.5
	) +
	geom_point(
		data = subset(all.lines, Tx.Type == "cdk" & Interim.Platinum == 1),
		aes(x = -10,
		    y = index),
		fill = "#00B26E",
		color = "black",
		shape = 22,
		size = 4.5
	) +
	geom_point(
		data = subset(all.lines, Tx.Type == "cdk" & Interim.Platinum == 0),
		aes(x = -10,
		    y = index),
		fill = "#B2E1C4",
		color = "black",
		shape = 22,
		size = 4.5
	) + 
	theme_classic() +
	theme(axis.text.y = element_blank(), 
	      axis.line.y = element_blank(), 
	      axis.ticks = element_blank(),
	      legend.position = "none")

print(plot_)

all.lines %>%
dplyr::select(index, Tx.Type, Tx.Line, deidentified_time, event, Reason.Stop) %>%
dplyr::rename(time = deidentified_time) %>%
dplyr::filter(Tx.Type %in% c("cdk", "parp")) %>%
dplyr::distinct() %>%
dplyr::mutate(event = case_when(is.na(event)~ 0, TRUE ~ event)) %>%
coxph(Surv(time,event) ~ Tx.Type + strata(index), data = .) %>%
summary()

all.lines %>%
dplyr::select(index, Tx.Type, Best.Response) %>%
dplyr::rename(response = Best.Response) %>%
dplyr::filter(!is.na(response)) %>%
dplyr::filter(Tx.Type %in% c("cdk", "parp")) %>%
dplyr::filter(response != "No_scan") %>%
dplyr::mutate(OR = case_when(response %in% c("PR", "CR") ~ 1, TRUE ~ 0)) %>%
dplyr::select(OR, Tx.Type) %>%
table() %>%
fisher.exact()
```

```{r pre_post_acquired_B2, eval = TRUE, echo = FALSE, message = FALSE, include = TRUE, fig.align = 'left', fig.height = 7, fig.width = 21, fig.cap = '**Figure 2g:**'}
all.pw2.plus = suppressWarnings(readr::read_csv(file = "../data/Supp_Fig2g_pre_post_B2_ANON.csv.gz", col_names = TRUE, col_types = cols(.default = col_character()))) %>%
	       readr::type_convert() %>%
	       dplyr::select(-X1) %>%
	       dplyr::rename(Patient_Anonymized = Anonymized.ID) %>%
 	       dplyr::mutate(across(Post_RB1:Post_CCNE1, ~factor(.x, levels = c("Not Tested", "WT", "Deletion", "Amplification", "Fusion", "Splice_driver", "Missense_driver", "Truncating_driver", "Amplification;acquired", "Deletion;acquired", "Fusion;acquired", "Splice_driver;acquired", "Missense_driver;acquired", "Truncating_driver;acquired")))) %>%
 	       dplyr::mutate(BRCA2 = factor(BRCA2, levels = c("WT", "GL"))) %>%
               dplyr::arrange(-as.numeric(BRCA2),
			      -as.numeric(Post_RB1),
			      -as.numeric(Post_PTEN),
			      -as.numeric(Post_AURKA),
			      -as.numeric(Post_LATS2),
			      -as.numeric(Post_YAP1),
			      -as.numeric(Post_FAT1),
			      -as.numeric(Post_CCNE1))

{get_type_fun = function(x) strsplit(x, ";")[[1]]

col_op = c("WT" = "#CCCCCC",
	   "GL" = brewer.pal(6, "BuPu")[6],
	   "SOM" = brewer.pal(6, "BuPu")[3],
	   "gWT" = brewer.pal(6, "BuPu")[1],
           "Missense_driver" =  "#005000",
	   "Splice_driver" = "#EE7600",
	   "Truncating_driver" = "#000000", 
	   "Fusion" = "purple", 
	   "Amplification" = "red", "Deletion" = "blue")

alter_fun = list(
    background = function(x,y,w,h) {grid.rect(x,y,w,h, gp = gpar(fill = "white", col = NA))}, 
   gWT = function(x, y, w, h) {
    grid.rect(x, y, w-unit(2, "pt"), h-unit(2, "pt"),
              gp = gpar(fill = col_op["gWT"], col = NA))
   },
   GL = function(x, y, w, h) {
    grid.rect(x, y, w*0.9, h*0.9,
              gp = gpar(fill = col_op["GL"], col = NA))
   },
   SOM= function(x, y, w, h) {
    grid.rect(x, y, w*0.9, h*0.9,
              gp = gpar(fill = col_op["SOM"], col = NA))
   },
   Amplification = function(x, y, w, h) {
    grid.rect(x, y, w*0.9, h*0.9,
              gp = gpar(fill = col_op["Amplification"], col = NA))
   },
    Deletion = function(x, y, w, h) {
    grid.rect(x, y, w*0.9, h*0.9,
              gp = gpar(fill = col_op["Deletion"], col = NA))
   },
   Missense_driver = function(x, y, w, h) {
    grid.rect(x, y, w*0.9, h*0.33,
              gp = gpar(fill = col_op["Missense_driver"], col = NA))
   },
   Splice_driver = function(x, y, w, h) {
    grid.rect(x, y, w*0.9, h*0.33,
              gp = gpar(fill = col_op["Splice_driver"], col = NA))
   },
   Truncating_driver = function(x, y, w, h) {
    grid.rect(x, y, w*0.9, h*0.33,
              gp = gpar(fill = col_op["Truncating_driver"], col = NA))
   },
   Fusion = function(x, y, w, h) {
     grid.rect(x, y, w-unit(2, "pt"), h*0.33,
              gp = gpar(fill = col_op["Fusion"], col = NA))
  },
   acquired = function(x, y, w, h) {
    grid.rect(x, y, w, h,
              gp = gpar(fill = NA, lwd = 2))
   },
   Not_Tested = function(x,y,w,h) {
     grid.points(x,y,pch = "-")
   },
     WT = function(x, y, w, h) {
    grid.rect(x, y, w*0.9, h*0.9,
              gp = gpar(fill = col_op["WT"], col = NA))
   }
)
  
heatmap_legend_param = list(title = "Variants (Annotated Drivers)",
			    at = c("GL", "SOM", "Missense_driver", "Splice_driver", "Truncating_driver", "Amplification", "Deletion", "acquired"), 
                            labels = c("gBRCA2", "sBRCA2", "Missense mutation (putative driver)", "Splice Site mutation (putative driver)", "Truncating mutation (putative driver)", "Amplification", "Homozygous deletion", "Acquired on CDK4/6"),
			    ncol = 3)

awp = all.pw2.plus %>%
      dplyr::select(Patient_Anonymized, BRCA2, Post_RB1, Post_PTEN, Post_AURKA, Post_LATS2, Post_YAP1, Post_FAT1, Post_CCNE1)

abt.plus = t(awp)
colnames(abt.plus) = abt.plus[1,]
abt.plus = abt.plus[-1,]
samp.order = colnames(abt.plus)

ann.op.plus = all.pw2.plus %>%
 	      dplyr::pull(BRCA2)
col.ann = list(BRCA2 = c("GL" = brewer.pal(6, "BuPu")[6], "gWT" = brewer.pal(6, "BuPu")[1]))

ht_opt$COLUMN_ANNO_PADDING = unit(8, "mm")
op_pre_post = oncoPrint(as.matrix(abt.plus),
			alter_fun = alter_fun,
			col = col_op,
			column_order = samp.order,
			row_names_side = "left",
			pct_side = "right",
			row_names_gp = gpar(fontsize = 12), 
			column_split = all.pw2.plus$BRCA2, column_gap = unit(6, "mm"),  
			row_gap = unit(3, "mm"),
			row_order = row.names(abt.plus),
			top_annotation = NULL,
			# bottom_annotation = annot.bottom,
			right_annotation = NULL,
			show_pct = FALSE, 
			column_title = "Pre-post CDK4/6 samples ", 
			heatmap_legend_param = heatmap_legend_param,
			alter_fun_is_vectorized = TRUE)
 
suppressWarnings(draw(op_pre_post, heatmap_legend_side = "bottom", annotation_legend_side = "bottom"))
}
```

<br/>

```{r, eval = TRUE, echo = FALSE, message = FALSE, include = TRUE}
b2.pw = all.pw2.plus %>%
 	filter(BRCA2 == "GL")
wt.pw = all.pw2.plus %>%
 	filter(BRCA2 == "WT")
b2.acq = length(grep("acquired", b2.pw$Post_RB1))
wt.acq = length(grep("acquired", wt.pw$Post_RB1))
fisher.exact(matrix(c(b2.acq , dim(b2.pw)[1] - b2.acq, wt.acq , dim(wt.pw)[1] - wt.acq), nrow = 2))
```
